Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-18 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

\begin{definition}[QIVPM]
  Assume a collection of intervals $\mathscr{I}$ including $\imposs$
  and $\necess$ with addition and scalar multiplication defined as
  follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [\ell_{1},r_{1}]+[\ell_{2},r_{2}]=[\ell_{1}+\ell_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[\ell,r]=\begin{cases}
  [x\ell,xr] & \textrm{for }x\ge0\,;\\{}
  [xr,x\ell] & \textrm{for }x\le0\,.
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  Then we take a QIVPM~$\bar{\mu}$ to be an assignment of an interval to each
  event (projection operator~$P$) subject to the following constraints:
  \begin{subequations}\label{eq:QIVPM-constraints}
  \begin{eqnarray}
   &  & \bar{\mu}(\mathbb{0})=\imposs\,,\\
   &  & \bar{\mu}(\mathbb{1})=\necess\,,\label{eq:necess}\\
   &  & \bar{\mu}\left(\mathbb{1}-P\right)= \necess -\bar{\mu}\left(P\right)\,,
  \end{eqnarray}
  \end{subequations}
  and satisfying for each pair of \emph{commuting} projectors~$P_0$
  and~$P_1$ with $P_0P_1=P_1P_0$,
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
classical IVPMs with quantum probability measures.

@@

In Sec.~\ref{sec:Interval-Uncertainty}, we discussed how states
consistent with a QIVPM~$\bar{\mu}$ on the projectors in a subspace
of event~$\events'$. In particular, if $\events'$ commutes, $\bar{\mu}$
restricting on $\events'$ is essentially a classical IVPM, and Thm.~\ref{thm:Shapley}
claims that there is always a state~$\rho$ consistent with $\bar{\mu}$.
In this section, we want to shift our attention to consider states
consistent with $\bar{\mu}$ on \emph{every} event, $\coreBorn\left(\bar{\mu},\events\right)$.
When the intervals are infinitely precise, i.e., $\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$ becomes
quantum probability measure~\citep{HOSTunpublished}. In this situation,
Gleason's theorem claims that there is always a unique state consistent
with $\bar{\mu}$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$,
the state consistent with it on every projector, $\coreBorn\left(\bar{\mu},\events\right)$,
is unique. \end{thm}

\noindent However, if we consider general QIVPMs mapping to an \emph{arbitrary}
set of intervals, there is a QIVPM that no state can be consistent
with it.

@@

\begin{definition}[{morphism}]{\label{def:morphism}}Given a collection
of intervals $\mathscr{I}$ including $\imposs$ and $\necess$, a
function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is called
a \emph{morphism}\yutsung{We can give it a better name. If sticking
with this name, we may need to prove that it is a morphism in the
sense of category theory or something...}if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent Because the similarity between Def.~\ref{def:morphism}
and Def.~\ref{def:QIVPM}, $\iota\circ\mu$ should always be a IVPM
no matter in classical or quantum case. Surprisingly, if $\iota\circ\mu$
is always a IVPM, then $\iota$ must be a morphism.

\begin{thm}\label{thm:iota-statements}The following three statements
are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}The function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is a morphism.
\item \label{enu:iota-mu-CIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a classical IVPM for any classical probability measure~$\mu:\events_{C}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

As we have discussed, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM}. The rest of the proof is to prove that if
$\iota$ is not a morphism, then $\iota\circ\mu$ might not be a QIVPM
for any quantum probability measure~$\mu$. If $\iota$ doesn't satisfy
Eq.~(\ref{eq:iota-Inclusion}), we can find three particular numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$ such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$,
but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a morphism if and only
if $\iota\circ\mu$ is always a QIVPM. Similarly, Statement~\ref{enu:iota-subject-to}
and \ref{enu:iota-mu-CIVPM} are equivalent as well~\citep{HOSTunpublished}.

Even if $\iota$ is not a morphism, the composited function~$\iota\circ\mu$
may still be a QIVPM under so suitable conditions. For example, Eq.~(\ref{eq:iota-Inclusion})
can be weaken as follow: for any numbers~$x_{0}$ and $x_{1}\in\left[0,1\right]$,
their sum~$x_{0}+x_{1}\in\left[0,1\right]$ implies
\begin{equation}
\iota\left(x_{0}+x_{1}\right)\subseteq\iota\left(x_{0}\right)+\iota\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
Then, given a Hilbert space of dimension~$d\le3$, when $\iota$
satisfies Eqs.~(\ref{eq:iota-constraints}) and Eq.~(\ref{eq:iota-add}),$\iota\circ\mu$
still must be a QIVPM~\citep{HOSTunpublished}.

@@

Since morphisms have many desirable properties, the intervals it mapping
to either is trivial or contains infinity intervals.

\begin{thm}\label{thm:convex-uncountable}For any function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
described in Thm.~\ref{thm:iota-statements}, either $\mathscr{I}=\mathscr{I}_{0}$
defined in Eq.(\ref{eq:3-value-intervals}) or $\mathscr{I}$ contains
uncountably many intervals.\end{thm}

>>>>

\begin{definition}[QIVPM]\label{def:QIVPM}
  Assume a collection of intervals $\mathscr{I}$ including $\imposs$
  and $\necess$ with addition and scalar multiplication defined as
  follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [\ell_{1},r_{1}]+[\ell_{2},r_{2}]=[\ell_{1}+\ell_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[\ell,r]=\begin{cases}
  [x\ell,xr] & \textrm{for }x\ge0\,;\\{}
  [xr,x\ell] & \textrm{for }x\le0\,.
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  Then we take a QIVPM~$\bar{\mu}$ to be an assignment of an interval to each
  event (projection operator~$P$) subject to the following constraints:
  \begin{subequations}\label{eq:QIVPM-constraints}
  \begin{eqnarray}
   &  & \bar{\mu}(\mathbb{0})=\imposs\,,\\
   &  & \bar{\mu}(\mathbb{1})=\necess\,,\label{eq:necess}\\
   &  & \bar{\mu}\left(\mathbb{1}-P\right)= \necess -\bar{\mu}\left(P\right)\,,
  \end{eqnarray}
  \end{subequations}
  and satisfying for each pair of \emph{commuting} projectors~$P_0$
  and~$P_1$ with $P_0P_1=P_1P_0$,
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
\noindent The first three constraints,
Eqs.~(\ref{eq:QIVPM-constraints}), are the direct counterpart of the
corresponding ones for classical IVPMs.  With the understanding that
the union of classical sets $E_1\cup E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of quantum projection
operators~\cite{Griffiths2003}, the last condition,
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}), is a
direct counterpart of the convexity condition of
Eq.~(\ref{eq:classicalconvex}). Thus our definition of QIVPMs merges
classical IVPMs with quantum probability measures.

@@

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. In contrast with
the conventional Born rule~\citep{Born1983,peres1995quantum,544199,Jaeger2007}
and Gleason's theorem~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}:

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
the state consistent with it on every projector, $\coreBorn\left(\bar{\mu},\events\right)$,
is unique. \end{thm}

\noindent We haven't discussed how to construct a QIVPM according
to a simple rule. Even if we have a QIVPM, there might not have any
states consistent with it.

@@

\begin{definition}[THOS rule]\label{def:THOS}Given a collection
of intervals $\mathscr{I}$ including $\imposs$ and $\necess$, a
function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is called
a THOS rule if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is
subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent Because the similarity between Def.~\ref{def:THOS} and
Def.~\ref{def:QIVPM}, $\iota\circ\mu$ should always be a IVPM no
matter in classical or quantum case. Surprisingly, if $\iota\circ\mu$
is always a IVPM, then $\iota$ must be a THOS rule.

\begin{thm}[The equivalence among THOS rules and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}The function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is a THOS rule.
\item \label{enu:iota-mu-CIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a classical IVPM for any classical probability measure~$\mu:\events_{C}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

As we have discussed, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM}. The rest of the proof is to prove that if
$\iota$ is not a THOS rule, then $\iota\circ\mu$ might not be a
QIVPM for any quantum probability measure~$\mu$. If $\iota$ doesn't
satisfy Eq.~(\ref{eq:iota-Inclusion}), we can find three particular
numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$ such
that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a THOS rule if and only
if $\iota\circ\mu$ is always a QIVPM. Similarly, Statement~\ref{enu:iota-subject-to}
and \ref{enu:iota-mu-CIVPM} are equivalent as well~\citep{HOSTunpublished}.

Even if $\iota$ is not a THOS rule, the composited function~$\iota\circ\mu$
may still be a QIVPM under so suitable conditions. For example, Eq.~(\ref{eq:iota-Inclusion})
can be weaken as follow: for any numbers~$x_{0}$ and $x_{1}\in\left[0,1\right]$,
their sum~$x_{0}+x_{1}\in\left[0,1\right]$ implies
\begin{equation}
\iota\left(x_{0}+x_{1}\right)\subseteq\iota\left(x_{0}\right)+\iota\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
Then, given a Hilbert space of dimension~$d\le3$, when $\iota$
satisfies Eqs.~(\ref{eq:iota-constraints}) and Eq.~(\ref{eq:iota-add}),$\iota\circ\mu$
still must be a QIVPM~\citep{HOSTunpublished}.

@@

Since THOS rules have many desirable properties, the intervals it
mapping to either is trivial or contains infinity intervals.

\begin{thm}[{The range of THOS rule}s]\label{thm:convex-uncountable}For
any function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ described
in Thm.~\ref{thm:iota-statements}, either $\mathscr{I}=\mathscr{I}_{0}$
defined in Eq.(\ref{eq:3-value-intervals}) or $\mathscr{I}$ contains
uncountably many intervals.\end{thm}
